Red, Blue, or Purple? 2024’s Battleground States Tell a Story

### 2024-11-04 paper/paper.qmd
paper correction and styling

library(tidyverse)
library(arrow)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(here)
library(knitr)
library(rstanarm)
library(modelsummary)

@@

# Load necessary libraries
library(tidyverse)
library(here)
library(readr)

@@

1. State Effects ($\alpha_i$)
```{r}
#| label: tbl-state_effects
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| 
library(dplyr)
library(readr)
library(here)
library(writexl) 
library(kableExtra)

# Read in the cleaned data
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)

# Calculate the average state effect based on polling percentage (pct) by state and pollster_rating_name
state_effects <- data %>%
  group_by(state, pollster_rating_name) %>%
  summarise(
    state_effect = round(mean(pct, na.rm = TRUE), 1)  # Average polling percentage for each state-pollster combination, rounded to 1 decimal
  )

# Display the state effects
#print(state_effects)

# Save state_effects to a CSV file
write_csv(state_effects, path = here("data", "02-analysis_data", "state effect.csv"))

# Read the saved CSV file
state_effects_read <- read_csv(here("data", "02-analysis_data", "state effect.csv"), show_col_types = FALSE)

# Display the first 5 rows as a table using kable
state_effects_read %>%
  head(5) %>%
  kable(
    col.names = c("State", "Pollster Rating Name", "State Effect"),
    caption = "State Effects Example",
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"))

@@

2. Pollster Effect ($\beta_j$)
```{r}
#| label: tbl-pollster_effect
#| echo: false
#| warning: false
#| message: false
# Load necessary libraries
library(dplyr)
library(readr)
library(here)

# Read in the cleaned data
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)

# Check if the column names are as expected
#print(colnames(data))

# Calculate the average pollster effect based on pollscore
pollster_effects <- data %>%
  group_by(pollster_rating_name) %>%
  summarise(
    pollster_effect = mean(pollscore, na.rm = TRUE)  # Calculate average pollscore for each pollster
  )

# Display the pollster effects
#print(pollster_effects)

# Save pollster effects to a CSV file, including the pollster_rating_name
write_csv(pollster_effects, path = here("data", "02-analysis_data", "pollster effect.csv"))

# Read the saved CSV file
pollster_effects_read <- read_csv(here("data", "02-analysis_data", "pollster effect.csv"), show_col_types = FALSE)

# Display the first 5 rows as a table using kable
pollster_effects_read %>%
  head(5) %>%
  kable(
    col.names = c("Pollster Rating Name", "Pollster Effect"),
    caption = "Pollster Effects Example",
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"))

@@

3. Candidate Fixed Effect ($\gamma_k$)

```{r}
#| label: tbl-candidate_fixed_effect
#| echo: false
#| warning: false
#| message: false
# Load necessary libraries
library(dplyr)
library(readr)
library(here)
library(knitr)
library(kableExtra)

# Read in the cleaned data
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)

# Calculate the candidate fixed effect as the average support across all polls
candidate_fixed_effect <- data %>%
  group_by(candidate_name) %>%
  summarise(gamma_k = round(mean(pct, na.rm = TRUE), 1))

# Display candidate fixed effects
#print(candidate_fixed_effect)

# Save candidate_fixed_effect to a CSV file
write_csv(candidate_fixed_effect, path = here("data", "02-analysis_data", "candidate_fixed_effect.csv"))

@@

# Display the first 5 rows as a table using kable
candidate_fixed_effect_read %>%
  head(5) %>%
  kable(
    col.names = c("Candidate Name", "Candidate Fixed Effect (gamma_k)"),
    caption = "Candidate Fixed Effects Example",
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"))

@@

4. Adjust predictions by poll score and sample size $( \delta \cdot \mathrm{pct}_{ijk} + \eta \cdot \mathrm{sample\_size}_{ijk} )$

```{r}
#| label: tbl-adjusted_prediction
#| echo: false
#| warning: false
#| message: false
# Load necessary libraries
library(dplyr)
library(readr)
library(here)
library(knitr)
library(kableExtra)

# Read in the cleaned data
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)

# Ensure the pollscore and sample_size columns are non-NA (replace NAs with default values if needed)
data <- data %>%
  mutate(
    pollscore = ifelse(is.na(pollscore), 0, pollscore),  # Replace NA pollscore with 0
    sample_size = ifelse(is.na(sample_size), 1, sample_size)  # Replace NA sample_size with 1
  )

# Calculate the weight for each poll as the product of pollscore and sample_size
data <- data %>%
  mutate(weight = pollscore * sample_size)

# Calculate the weighted average of pct for each candidate in each state
weighted_predictions <- data %>%
  group_by(state, candidate_name) %>%
  summarise(
    weighted_pct = round(sum(pct * weight, na.rm = TRUE) / sum(weight, na.rm = TRUE), 0),
    total_weight = round(sum(weight, na.rm = TRUE), 0)  # Optional: To inspect total weight per group, rounded to 0 dp
  ) %>%
  ungroup()

# Display the weighted predictions
#print(weighted_predictions)

# Save weighted_predictions to an csv file
write_csv(weighted_predictions, path = here("data", "02-analysis_data", "weighted_predictions.csv"))

# Optionally, display the first 5 rows as a table using kable
weighted_predictions %>%
  head(5) %>%
  kable(
    col.names = c("State", "Candidate Name", "Weighted Percentage", "Total Weight"),
    caption = "Weighted Predictions by Poll Score and Sample Size Example",
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down", "hold_position"))

@@

5. Adjustment of weighted predictions by accounting for recency $( \theta \cdot \mathrm{days\_to\_election}_{ijk} )$

```{r}
#| label: tbl-weighted_prediction
#| echo: false
#| warning: false
#| message: false
# Load necessary libraries
library(dplyr)
library(readr)
library(here)
library(knitr)
library(kableExtra)

# Read in the cleaned data
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)

# Ensure no NA values in pollscore, sample_size, and days_to_election
data <- data %>%
  mutate(
    pollscore = ifelse(is.na(pollscore), 0, pollscore),  
    sample_size = ifelse(is.na(sample_size), 1, sample_size),  
    days_to_election = ifelse(is.na(days_to_election) | days_to_election <= 0, 
                               max(days_to_election, na.rm = TRUE), 
                               days_to_election)  
  )

# Calculate recency weight
data <- data %>%
  mutate(recency_weight = ifelse(days_to_election > 0, 1 / days_to_election, 0))

# Calculate total weight
data <- data %>%
  mutate(total_weight = pollscore * sample_size * recency_weight)

# Calculate the weighted average of pct for each candidate in each state
weighted_predictions <- data %>%
  group_by(state, candidate_name) %>%
  summarise(
    weighted_pct = round(ifelse(sum(total_weight, na.rm = TRUE) == 0, 0, 
                                sum(pct * total_weight, na.rm = TRUE) / sum(total_weight, na.rm = TRUE)), 1),
    total_weight = round(sum(total_weight, na.rm = TRUE), 1)
  ) %>%
  ungroup()

# Save weighted predictions to csv
write_csv(weighted_predictions, path = here("data", "02-analysis_data", "weighted_predictions_with_recency.csv"))

# Display first 5 rows without extra styling
weighted_predictions %>%
  head(5) %>%
  kable(
    col.names = c("State", "Candidate Name", "Weighted Percentage", "Total Weight"),
    caption = "Weighted Predictions with Recency Adjustment Examples"
  )
  # Avoid using `booktabs` or `latex_options` here to prevent LaTeX conflicts

# Pollster reliability adjustment
zeta <- 0.2  
data <- data %>%
  mutate(
    pollster_reliability_adjustment = round(zeta * pollscore, 1)  # Rounded to 1 decimal place
  )

# Save pollster reliability adjustments to csv
write_csv(data, path = here("data", "02-analysis_data", "Pollscore_reliability_adjust.csv"))
```

@@

6. Pollscore Reliability ($\zeta$)
```{r}
#| echo: false
#| warning: false
#| message: false
# Read in the cleaned data
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)

@@

```{r}
#| label: tbl-modelresults
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(readr)
library(here)
library(knitr)
library(kableExtra)

# Load the main data and effect files
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)
pollster_effect <- read_csv(here("data", "02-analysis_data", "pollster effect.csv"), show_col_types = FALSE)
state_effect <- read_csv(here("data", "02-analysis_data", "state effect.csv"), show_col_types = FALSE)
candidate_fixed_effect <- read_csv(here("data", "02-analysis_data", "candidate_fixed_effect.csv"), show_col_types = FALSE)
polling_percentage_sample_size_adjust <- read_csv(here("data", "02-analysis_data", "Adjust predictions by poll score and sample size.csv"), show_col_types = FALSE)
days_to_election_adjust <- read_csv(here("data", "02-analysis_data", "adjust weights for recency by using days_to_election.csv"), show_col_types = FALSE)
Pollscore_reliability_adjust <- read_csv(here("data", "02-analysis_data", "Pollscore_reliability_adjust.csv"), show_col_types = FALSE)

@@

# Display the results in a kable table with kableExtra styling
results %>%
  kable(caption = "Final Predictions for Trump and Harris") %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 14)

@@

```{r}
#| label: fig-presidential-pies
#| fig-cap: "2024 Presidential Predictions"
#| echo: false
#| warning: false
#| message: false
#| fig-width: 40
#| fig-height: 15
#| fig.fullwidth: true

# Load required libraries
library(ggplot2)
library(gridExtra)
library(grid)

@@

# Define colors for Trump (red), Harris (blue), and Others (gray)
colors <- c("Trump" = "red", "Harris" = "blue", "Others" = "gray80")

# First Pie Chart: Trump Prediction
p1 <- ggplot(data_trump, aes(x = "", y = percentage, fill = candidate)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = sprintf("%.2f%%", percentage)), 
            position = position_stack(vjust = 0.5), 
            size = 11, fontface = "bold") +
  scale_fill_manual(values = colors) +
  theme_void() +
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

# Second Pie Chart: Harris Prediction
p2 <- ggplot(data_harris, aes(x = "", y = percentage, fill = candidate)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = sprintf("%.2f%%", percentage)), 
            position = position_stack(vjust = 0.5), 
            size = 11, fontface = "bold") +
  scale_fill_manual(values = colors) +
  theme_void() +
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

# Third Pie Chart: Trump vs. Harris vs. Others
p3 <- ggplot(data_both, aes(x = "", y = percentage, fill = candidate)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = sprintf("%.2f%%", percentage)), 
            position = position_stack(vjust = 0.5), 
            size = 11, fontface = "bold") +
  scale_fill_manual(values = colors) +
  theme_void() +
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

# Create labels for each chart with increased font size
label_a <- textGrob("(a) Trump Vote Share", gp = gpar(fontsize = 35, fontface = "bold"))  # Increased font size
label_b <- textGrob("(b) Harris Vote Share", gp = gpar(fontsize = 35, fontface = "bold"))  # Increased font size
label_c <- textGrob("(c) Comparative Vote Share", gp = gpar(fontsize = 35, fontface = "bold"))  # Increased font size

# Arrange the pie charts in a single row with labels below each plot
grid.arrange(
  p1, p2, p3,
  label_a, label_b, label_c,
  ncol = 3,
  layout_matrix = rbind(c(1, 2, 3), c(4, 5, 6)),  # Arrange plots in the first row and labels in the second
  heights = c(10, 1.5)  # Allocate more space to the label row to avoid compression
)

@@

The first pie chart, labeled (a) Trump Vote Share, displays Donald Trump’s projected support level if we consider his vote share against all other potential candidates. According to the model, Trump is estimated to hold 46.11% of the vote, while 53.89% falls under the "Others" category. This result suggests that although Trump has a strong share of the projected votes, he does not exceed the 50% threshold in isolation, indicating there remains a sizable portion of voters who may support other candidates or remain undecided.

The second pie chart, (b) Harris Vote Share, presents Kamala Harris’s estimated support level against all other candidates. Harris is projected to receive 45.65% of the vote, with 54.35% categorized as "Others." Similar to Trump’s results, Harris shows a solid portion of support, though she too falls short of the majority. This chart highlights that Harris’s vote share is almost on par with Trump’s but leaves a considerable portion of votes uncommitted to either of the main candidates.

The third pie chart, (c) Comparative Vote Share, illustrates the direct comparison between Trump, Harris, and the "Others" category. In this model, Trump holds a slight lead with 46.11%, while Harris closely follows with 45.65%. Meanwhile, 8.24% of the vote remains attributed to "Others," representing voters who may support third-party candidates or are undecided. This small lead for Trump, amounting to just a 0.46% difference, emphasizes the closeness of the race. The presence of a significant "Others" portion could prove crucial if those votes shift toward one of the primary candidates, potentially altering the predicted outcome.

@@

```{r}
#| label: tbl-modelresults_states
#| echo: false
#| warning: false
#| message: false
library(dplyr)

@@

```{r}
#| label: fig-modelresults_states
#| echo: false
#| warning: false
#| message: false
# Load necessary libraries
library(tidyverse)

@@

```{r}
#| label: fig-modelresults_map
#| echo: false
#| warning: false
#| message: false
#| fig-width: 20
#| fig-height: 8
#| fig.fullwidth: true
# Load necessary libraries
library(tidyverse)
library(readxl)
library(ggplot2)
library(usmap)  # For easier mapping of U.S. states
library(here)

@@

# Load necessary libraries
library(tidyverse)
library(usmap)
library(ggplot2)
library(here)

# Load the state effect data
state_effect <- read_csv(here("data", "02-analysis_data", "state effect.csv"))

@@

# Load necessary libraries
library(tidyverse)
library(here)

# Load the data (adjust the file path as needed)
pollster_effects <- read_csv(here("data", "02-analysis_data", "pollster effect.csv"))

@@

# Load necessary libraries
library(tidyverse)
library(here)

@@

```{r}
# Load required libraries
library(readr)
library(here)
library(dplyr)
library(ggplot2)
library(gridExtra) # For arranging plots side by side

@@

# Plot for Donald Trump
plot_trump <- ggplot(trump_data, aes(x = source, y = weighted_pct, fill = source)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.shape = NA) +
  labs(title = "Donald Trump", y = "Weighted Prediction (%)") +
  theme_minimal() +
  theme(axis.title.x = element_blank()) # Remove x-axis title

# Plot for Kamala Harris
plot_harris <- ggplot(harris_data, aes(x = source, y = weighted_pct, fill = source)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.shape = NA) +
  labs(title = "Kamala Harris", y = "Weighted Prediction (%)") +
  theme_minimal() +
  theme(axis.title.x = element_blank()) # Remove x-axis title

# Arrange the two plots side by side
grid.arrange(plot_trump, plot_harris, ncol = 2)

>>>>

library(tidyverse)
library(arrow)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(here)
library(knitr)
library(rstanarm)
library(modelsummary)
library(readr)
library(writexl) 
library(ggplot2)
library(gridExtra)
library(grid)
library(usmap)

@@

@@

#### State Effects ($\alpha_i$)
```{r}
#| label: tbl-state_effects
#| tbl-cap: "State Effects Example"
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Read the saved CSV file
state_effects_read <- read_csv(here("data", "02-analysis_data", "state_effect.csv"), show_col_types = FALSE)

# Display the first 5 rows as a table using kable
state_effects_read %>%
  head(5) %>%
  kable(
    col.names = c("State", "Pollster Rating Name", "State Effect"),
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped"))

@@

#### Pollster Effect ($\beta_j$)
```{r}
#| label: tbl-pollster_effect
#| tbl-cap: "Pollster Effects Example"
#| echo: false
#| warning: false
#| message: false

# Read the saved CSV file
pollster_effects_read <- read_csv(here("data", "02-analysis_data", "pollster_effect.csv"), show_col_types = FALSE)

# Display the first 5 rows as a table using kable
pollster_effects_read %>%
  head(5) %>%
  kable(
    col.names = c("Pollster Rating Name", "Pollster Effect"),
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped"))

@@

#### Candidate Fixed Effect ($\gamma_k$)

```{r}
#| label: tbl-candidate_fixed_effect
#| tbl-cap: "Candidate Fixed Effects Example"
#| echo: false
#| warning: false
#| message: false

@@

# Display the first 5 rows as a table using kable
candidate_fixed_effect_read %>%
  head(5) %>%
  kable(
    col.names = c("Candidate Name", "Candidate Fixed Effect (gamma_k)"),
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped"))

@@

#### Adjust predictions by poll score and sample size $( \delta \cdot \mathrm{pct}_{ijk} + \eta \cdot \mathrm{sample\_size}_{ijk} )$

```{r}
#| label: tbl-adjusted_prediction
#| tbl-cap: "Weighted Predictions by Poll Score and Sample Size Example"
#| echo: false
#| warning: false
#| message: false

# Read the saved CSV file
weighted_predictions <- read_csv(here("data", "02-analysis_data", "weighted_predictions.csv"), show_col_types = FALSE)

weighted_predictions %>%
  head(5) %>%
  kable(
    col.names = c("State", "Candidate Name", "Weighted Percentage", "Total Weight"),
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped"))

@@

#### Adjustment of weighted predictions by accounting for recency $( \theta \cdot \mathrm{days\_to\_election}_{ijk} )$

```{r}
#| label: tbl-weighted_prediction
#| tbl-cap: "Weighted Predictions with Recency Adjustment Examples"
#| echo: false
#| warning: false
#| message: false

# Read the saved CSV file
weighted_predictions <- read_csv(here("data", "02-analysis_data", "weighted_predictions_with_recency.csv"), show_col_types = FALSE)

# Display first 5 rows without extra styling
weighted_predictions %>%
  head(5) %>%
  kable(
    col.names = c("State", "Candidate Name", "Weighted Percentage", "Total Weight"),
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("striped"))
```

@@

#### Pollscore Reliability ($\zeta$)
```{r}
#| echo: false
#| warning: false
#| message: false
# Read in the cleaned data
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)

@@

```{r}
#| label: tbl-modelresults
#| tbl-cap: "Final Predictions for Trump and Harris"
#| echo: false
#| warning: false
#| message: false

# Load the main data and effect files
data <- read_csv(here("data", "02-analysis_data", "clean_data.csv"), show_col_types = FALSE)
pollster_effect <- read_csv(here("data", "02-analysis_data", "pollster_effect.csv"), show_col_types = FALSE)
state_effect <- read_csv(here("data", "02-analysis_data", "state_effect.csv"), show_col_types = FALSE)
candidate_fixed_effect <- read_csv(here("data", "02-analysis_data", "candidate_fixed_effect.csv"), show_col_types = FALSE)
polling_percentage_sample_size_adjust <- read_csv(here("data", "02-analysis_data", "adjust_predictions_by_poll_score_and_sample_size.csv"), show_col_types = FALSE)
days_to_election_adjust <- read_csv(here("data", "02-analysis_data", "adjust_weights_for_recency_by_using_days_to_election.csv"), show_col_types = FALSE)
Pollscore_reliability_adjust <- read_csv(here("data", "02-analysis_data", "pollscore_reliability_adjust.csv"), show_col_types = FALSE)

@@

# Display the results in a kable table with kableExtra styling
results %>%
  kable() %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 14)

@@

```{r}
#| label: fig-presidential-pies
#| fig-cap: "2024 Presidential Predictions"
#| fig-subcap: ["Trump", "Harris", "Overall"]
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 3
#| layout-nrow: 1

@@

# Define colors for Trump (red), Harris (blue), and Others (gray)
colors <- c("Trump" = "red", "Harris" = "blue", "Others" = "darkgray")

# First Pie Chart: Trump Prediction
ggplot(data_trump, aes(x = "", y = percentage, fill = candidate)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = sprintf("%.2f%%", percentage)), 
            position = position_stack(vjust = 0.5), 
            size = 8, fontface = "bold", color = "white") +
  scale_fill_manual(values = colors) +
  theme_void() +
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

# Second Pie Chart: Harris Prediction
ggplot(data_harris, aes(x = "", y = percentage, fill = candidate)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = sprintf("%.2f%%", percentage)), 
            position = position_stack(vjust = 0.5), 
            size = 8, fontface = "bold", color = "white") +
  scale_fill_manual(values = colors) +
  theme_void() +
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

# Third Pie Chart: Trump vs. Harris vs. Others
ggplot(data_both, aes(x = "", y = percentage, fill = candidate)) +
  geom_bar(stat = "identity", width = 1, color = "black") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = sprintf("%.2f%%", percentage)), 
            position = position_stack(vjust = 0.5), 
            size = 8, fontface = "bold", color = "white") +
  scale_fill_manual(values = colors) +
  theme_void() +
  theme(legend.position = "right",
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        plot.margin = unit(c(0, 0, 0, 0), "cm"))

@@

The @fig-presidential-pies-1, displays Donald Trump’s projected support level if we consider his vote share against all other potential candidates. According to the model, Trump is estimated to hold 46.11% of the vote, while 53.89% falls under the "Others" category. This result suggests that although Trump has a strong share of the projected votes, he does not exceed the 50% threshold in isolation, indicating there remains a sizable portion of voters who may support other candidates or remain undecided.

The @fig-presidential-pies-2, presents Kamala Harris’s estimated support level against all other candidates. Harris is projected to receive 45.65% of the vote, with 54.35% categorized as "Others." Similar to Trump’s results, Harris shows a solid portion of support, though she too falls short of the majority. This chart highlights that Harris’s vote share is almost on par with Trump’s but leaves a considerable portion of votes uncommitted to either of the main candidates.

The @fig-presidential-pies-3, illustrates the direct comparison between Trump, Harris, and the "Others" category. In this model, Trump holds a slight lead with 46.11%, while Harris closely follows with 45.65%. Meanwhile, 8.24% of the vote remains attributed to "Others," representing voters who may support third-party candidates or are undecided. This small lead for Trump, amounting to just a 0.46% difference, emphasizes the closeness of the race. The presence of a significant "Others" portion could prove crucial if those votes shift toward one of the primary candidates, potentially altering the predicted outcome.

@@

```{r}
#| label: tbl-modelresults_states
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-modelresults_states
#| echo: false
#| warning: false
#| message: false

@@

```{r}
#| label: fig-modelresults_map
#| echo: false
#| warning: false
#| message: false
#| fig-width: 20
#| fig-height: 8
#| fig.fullwidth: true

@@

# Load the state effect data
state_effect <- read_csv(here("data", "02-analysis_data", "state_effect.csv"))

@@

# Load the data (adjust the file path as needed)
pollster_effects <- read_csv(here("data", "02-analysis_data", "pollster_effect.csv"))

@@

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| layout-ncol: 2
#| layout-nrow: 1

@@

# Plot for Donald Trump
ggplot(trump_data, aes(x = source, y = weighted_pct, fill = source)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.shape = NA) +
  labs(title = "Donald Trump", y = "Weighted Prediction (%)") +
  theme_minimal() +
  theme(axis.title.x = element_blank()) # Remove x-axis title

# Plot for Kamala Harris
ggplot(harris_data, aes(x = source, y = weighted_pct, fill = source)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.1, color = "black", fill = "white", outlier.shape = NA) +
  labs(title = "Kamala Harris", y = "Weighted Prediction (%)") +
  theme_minimal() +
  theme(axis.title.x = element_blank()) # Remove x-axis title
